---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rsketball

## Analysis of NBA players from 2001/02 to 2018/19 seasons using ESPN NBA

This package is designated for all NBA enthusiasts! This package works to scrape online tabular data from ESPN NBA website into a csv file. It also includes various functions to create graphs and statistical analysis for your interest (such as boxplots, player rankings by stats, and a summary statistics table).  

An example of the ESPN NBA 2018/19 Regular season player stats can be found in the following url:

https://www.espn.com/nba/stats/player/_/season/2019/seasontype/2

## Functions

`nba_scraper`
  
  * Scrapes data from ESPN NBA data into a csv file. User can specify the year of the season
  (2016, 2017, etc) and the season type (regular or playoffs).

`nba_boxplot`

   * Creates a boxplot of the categorical variable of interest on the y-axis and 
   the stat of interest on the x-axis.

`nba_ranking`
  
  * Generates a ranking and a visualization based on a column of a dataset  

`nbastats`

  * Generate summary stats for NBA players.
    The function filters the dataset further using the arguments provided and
    produces a tibble with summary statistics for a list of columns of a few players or teams.

## R Ecosystem

This 'rsketball' package aims to further gain understanding of ESPN NBA data and does not have a specific fit to the R ecosystem. There are currently some other library packages such as [`nbastatR`](https://www.rdocumentation.org/packages/nbastatR/versions/0.1.10131) that take data from other sources (NBA Stats API, Basketball Insiders, Basketball-Reference, HoopsHype, and RealGM), but no package that we currently know of takes data from ESPN NBA specifically.

### Installation

`rsketball` is still in project development. We estimate that by end March 2020, one can install the released version of `rsketball` from [CRAN](https://cran.r-project.org/).

Package installation in R:

```{r, eval=FALSE}
install.packages("rsketball")
```

And the development version from [Github](https://github.com/) with:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("UBC-MDS/rsketball")
```
