---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rsketball

## Analysis of NBA players from 2001/02 to 2018/19 seasons using ESPN NBA

This package is designated for all NBA enthusiasts! The `rsketball` package works to scrape online tabular data from the ESPN NBA website into a csv file. It also includes various functions to create graphs and statistical analysis for your interest (such as boxplots, player rankings by stats, and a summary statistics table).  

An example of the ESPN NBA 2018/19 Regular season player stats can be found in the following url:

https://www.espn.com/nba/stats/player/_/season/2019/seasontype/2

## Functions

`nba_scraper`
  
  * Scrapes data from ESPN NBA data into a csv file. User can specify the year of the season
  (2001 to 2019 at the moment) and the season type ("regular" or "postseason").

`nba_boxplot`

   * Creates a boxplot of the categorical variable of interest on the y-axis and 
   the stat of interest on the x-axis.

`nba_ranking`
  
  * Generates a ranking and a visualization based on a column of a dataset  

`nbastats`

  * Generate summary stats for NBA players.
    The function filters the dataset further using the arguments provided and
    produces a tibble with summary statistics for a list of columns of a few players or teams.

## Preparation Usage for Scraping with `nba_scraper()`

The `rsketball::nba_scraper` is based on Selenium (or specifically RSelenium) which enables automated web browsing through "drivers". __To use it, please ensure that `Docker` is installed.__

For installation instructions, please follow the [guide to Docker installation based on your OS type](https://ubc-mds.github.io/resources_pages/installation_instructions/). Docker will be used to pull relevant images that when executed as containers, will serve as the "drivers" for Selenium. 

The following steps are required only for the `nba_scraper` function. If you already have the scraped data file and wish to use the other functions (`nba_boxplot`, `nba_rank`, `nbastats`), there is no need to proceed with these steps.

__Step 1 (Command line/Terminal): Preparation Step (Docker container)__

Pull docker image with the following code in Terminal. We will stick to Chrome since it seems compatible with Windows while Firefox is not.
```sh
docker pull selenium/standalone-chrome
```
__Critical step about setting ports and memory allocation:__ 

We need to set up the Docker container default port 4444 to our computer host port 4445. Keep this port number as inputs for the `nba_scraper` function. We will also allocate the virtual memory of the container to 2Gb for it to scrape effectively. 

Run the following code in Terminal:
```sh
docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome
```
Verify that the docker container is in operation by running the following code in Terminal:
```sh
docker ps 
```
__Step 2 (R/RStudio): Scraping with `nba_scraper`__

Now that the container is running with the allocated memory and assigned port, we can proceed with testing

```R
library(rsketball)
# Scrape postseason season 2017/18 using "chrome" driver while saving to a local csv file.
nba_2017_playoffs <- nba_scraper(season_year = 2017, 
                                 season_type = "postseason",
                                 port=4445L, 
                                 sel_browser = "chrome",
                                 csv_path = "nba_2017_playoffs.csv")
```
If everything was executed as intended, you should obtain a csv file called "nba_2017_playoffs.csv" containing the scraped data, and a tibble in your R environment named "nba_2017_playoffs". With the tibble, you can use the other `rsketball` functions for your analysis.

__Step 3 (Command line/Terminal): Termination of Docker Container__

After test scraping is completed, we can shut down the Docker Container instance. This will also ensure that your computer memory/resources are restored.
```sh
docker stop $(docker ps -q)
```
If you wish to, you can also remove the Docker image from your computer, where "<image_id>" represents the id of your Docker image.
```sh
docker image rm <image_id>
```

## R Ecosystem

This `rsketball` package aims to further gain understanding of ESPN NBA data and does not have a specific fit to the R ecosystem. There are currently some other library packages such as [`nbastatR`](https://www.rdocumentation.org/packages/nbastatR/versions/0.1.10131) that take data from other sources (NBA Stats API, Basketball Insiders, Basketball-Reference, HoopsHype, and RealGM), but no package that we currently know of takes data from ESPN NBA specifically.

### Installation

`rsketball` is still in project development. We estimate that by end March 2020, one can install the released version of `rsketball` from [CRAN](https://cran.r-project.org/).

Package installation in R:

```{r, eval=FALSE}
install.packages("rsketball")
```

And the development version from [Github](https://github.com/) with:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("UBC-MDS/rsketball")
```
