---
title: "rsketball-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Analysis of NBA players from 2001/02 to 2018/19 seasons using ESPN NBA

This package is designated for all NBA enthusiasts! The `rsketball` package works to scrape online tabular data from the ESPN NBA website into a csv file. It also includes various functions to create graphs and statistical analysis for your interest (such as boxplots, player rankings by stats, and a summary statistics table).  

An example of the ESPN NBA 2018/19 Regular season player stats can be found in this [EPSN NBA url](https://www.espn.com/nba/stats/player/_/season/2019/seasontype/2)

## Functions

`nba_scraper`
  
  * Scrapes data from ESPN NBA data into a tibble, and allows user to store it into a csv file. User can specify the year of the season (2001 to 2019 at the moment) and the season type ("regular" or "postseason").

`nba_boxplot`

   * Creates a boxplot of the categorical variable of interest (TEAM or POS) on the y-axis and the numerical statistic of interest on the x-axis.

`nba_ranking`
  
  * Generates a ranking visualization based on the numerical statistic of interest column of a dataset.

`nba_team_stats`

  * Generate summary stats for NBA players. The function provides descriptive (mean, median, 25%, and 75% quantiles) team statistics of NBA data. It also comes with added functionalities for player position grouping per team too.

### Installation

`rsketball` is still in project development. We estimate that by end March 2020, one can install the released version of `rsketball` from [CRAN](https://cran.r-project.org/).

Package installation in R:

```{r, eval=FALSE}
install.packages("rsketball")
```

And the development version from [Github](https://github.com/) with:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("UBC-MDS/rsketball")
```

## Preparation Usage for Scraping with `nba_scraper()`

The `rsketball::nba_scraper` is based on Selenium (or specifically RSelenium) which enables automated web browsing through "drivers". __To use it, please ensure that `Docker` is installed.__

For installation instructions, please follow the [guide to Docker installation based on your OS type](https://ubc-mds.github.io/resources_pages/installation_instructions/). Docker will be used to pull the relevant Chromedriver image that when executed as containers, will serve as the "driver" for Selenium. 

The following steps are required only for the `nba_scraper` function. If you already have the scraped data file and wish to use the other functions (`nba_boxplot`, `nba_rank`, `nbastats`), there is no need to proceed with these steps.

__Step 1 (Command line/Terminal): Preparation of Docker container__

Pull the docker image with the following code in Terminal. We will stick to Chrome since it seems compatible with Windows while Firefox is not.
```sh
docker pull selenium/standalone-chrome
```
__Critical step about setting ports and memory allocation:__ 

We need to set up the Docker container default port 4444 to our computer host port 4445. Keep this port number as inputs for the `nba_scraper` function. We will also allocate 2GB of virtual memory for the container to scrape effectively. 

Run the following code in Terminal:
```sh
docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome
```
Verify that the docker container is in operation by running the following code in Terminal:
```sh
docker ps 
```
__Step 2 (R/RStudio): Scraping with `nba_scraper`__

Now that the container is running with the allocated memory and assigned port, we can proceed with testing

```R
library(rsketball)
library(RSelenium)
library(rvest)
library(tidyverse)
# Scrape postseason season 2017/18 while saving to a local csv file.
nba_2017_playoffs <- nba_scraper(season_year = 2017, 
                                 season_type = "postseason",
                                 port=4445L,
                                 csv_path = "nba_2017_playoffs.csv")
```
If everything was executed as intended, you should obtain a csv file called "nba_2017_playoffs.csv" containing the scraped data, and a tibble in your R environment named "nba_2017_playoffs". With the tibble, you can use the other `rsketball` functions for your analysis.

__Step 3 (Command line/Terminal): Termination of Docker Container__

After test scraping is completed, we can shut down the Docker Container instance. This will also ensure that your computer memory/resources are restored.
```sh
docker stop $(docker ps -q)
```
If you wish to, you can also remove the Docker image from your computer, where "<image_id>" represents the id of your Docker image.
```sh
docker image rm <image_id>
```

## Examples/Usage

To load the package:
```{r, message=FALSE}
library(rsketball)
```

### `nba_scraper`

nba_scraper() will help you create the dataframe of the NBA season of interest to conduct further analysis using the functions below. Here's a breakdown of the inputs:

- _season_year_: Numeric integer input between 2001 to 2019 (currently).

- _season_type_: String input of either "regular" or "postseason"

- _port_: Port number for matching Docker container port with Selenium remote driver instance. Defaults to 4445L. Positive integer that ends with L.

- _csv_path_: String input specifying output path of csv. Must end in ".csv". IF NULL, no csv file will be created. Defaults to NULL.

Below are two examples of the function for scraping:

Scrape regular season 2018/19 while saving to a csv file. This should return a tibble.

```{r, message=FALSE, eval = FALSE}
nba_2018 <- nba_scraper(season_year = 2018, 
                        season_type = "regular", 
                        port=4445L, 
                        csv_path = "nba_2018_regular.csv")
```

Scrape playoffs season 2017/18 without saving to a local csv file. This should return a tibble.

```{r, message=FALSE, eval=FALSE}
nba_2017 <- nba_scraper(season_year = 2017, 
                        season_type = "postseason",
                        port=4445L)
```

For the illustration of the other functions, let's create a toy dataset with similar properties as the scraped data from ESPN NBA.

```{r}
nba_data <- tibble::tibble(NAME = c("James", "Steph", "Bosh", "Klay", "Kobe"),
                    TEAM = c("MIA","GS","MIA","GS","LAL"),
                    POS = c("SF", "PG", "C", "SG", "SG"),
                    PTS = c(5,4,3,2,10),
                    TO = c(1,2,3,4,3),
                    `3PA` = c(10, 20, 30, 40, 50),
                    `FT%` = c(50, 60, 70, 80, 90))

```

### `nba_boxplot`

To further analyze how different teams and positions affect different scoring statistics, you can use `nba_boxplot()`. Here's a breakdown of the inputs:

- _nba_data_: Tibble of scraped nba data

- _team_or_position_: String input of either "team" or "position"

- _grouping_list_: Specific list inputs for grouping. Shows all values (of either "team" or "position") when NULL.

- _stats_column_: Numeric column name of interest. Not in string input format.

To look at the distribution of Points per player or 'PTS' (which is a numerical column) for different positions or 'POS' (which is a categorical column). 

Since grouping_list inputs are NULL, it will display the values of all possible positions.

```{r}
# Boxplot of all player positions based on PTS
nba_boxplot(nba_data = nba_data, 
            team_or_position = "position",
            grouping_list = NULL, 
            stats_column = PTS)
```

To look at the distribution of Three Pointers Attempts or '3PA' (which is a numerical column) for specific positions (must pass in a list). 

__Important:__ Since the column id starts with a number 3, we must ensure that the input for stats_column is formatted with backticks as shown:

```{r}
# Boxplot of specific player positions based on 3P%
nba_boxplot(nba_data, 
            team_or_position = "position",
            grouping_list = c("C","PG","SG"), 
            stats_column = `3PA`) # Formatted with backticks.
```

To look at the distribution of Free Throws Percentage or 'FT%' (which is a numerical column) for specific teams (must pass in a list). 

__Important:__ Since the column id has a "%" character in it, we must ensure that the input for stats_column is formatted with backticks as shown:

```{r}
nba_boxplot(nba_data, 
            team_or_position= "team", 
            grouping_list = c("MIA", "GS"), 
            stats_column = `FT%`) # Formatted with backticks.
```

###`nba_ranking`

The `nba_ranking()` function creates a visualization showing the rankings of a category with a statistic of interest. Here's a breakdown of its inputs:

- _nba_data_: Tibble of scraped nba data

- _column_: Categorical column name for ranking. Either NAME, TEAM or POS.

- _by_: Numeric statistic column name of interest. Not in string input format.

- _top_: Numeric limit of the top values for ranking. Defaults to 5

- _descending_: Boolean input to determine if ranking is in asc/descending order. Defaults to TRUE.

- _FUN_: Function for aggregation ranking. Defaults to the mean function.


In the first example, we ranked the top 3 players (NAME) based on their number of Three Points Attempts (3PA) made in a descending order.

__Important:__ Since the column id starts with a number 3, we must ensure that the input for stats_column is formatted with backticks as shown:

```{r}
# Find top 3 players for 3 Point Attempts (3PA) where higher is better
nba_ranking(nba_data,
            column = NAME,
            by = `3PA`, # Formatted with backticks.
            top = 3,
            descending = TRUE,
            FUN = mean)
```

In the second example, we ranked the top 3 players (NAME) based on their number of TurnOvers (TO) made in an ascending order.

```{r}
# Find top 2 teams for turnover (TO) where lower is better
nba_ranking(nba_data,
            column = TEAM,
            by = TO,
            top = 2,
            descending = FALSE,
            FUN = mean)
```

###`nba_team_stats`

The `nba_team_stats()` function finds statistics of mean, median, 25%, and 75% quantiles. This function is primarily focused on team, and allows for further grouping by player position per team. The function has the following inputs:

- _nba_data_: Tibble of scraped nba data

- _stats_filter_: List inputs of the different columns of interest. When NULL, it will show descriptive statistics for all numeric columns. Defaults to NULL.

- _teams_filter_: List inputs of the different teams of interest. When NULL, it will show the descriptive statistics for all available teams. Defaults to NULL.

- _positions_filter_: List inputs of the different player positions for each team. When NULL, it will not show any POS grouping. Defaults to NULL.

In the first example, we obtain the descriptive statistics for all available teams with the default NULL arguments in stats_filter, teams_filter, positions_filter.

```{r}
# Find descriptive stats for all teams without position
nba_team_stats(nba_data)
```

In the second example, we obtain the descriptive statistics of relevant numeric columns (PTS and TO) for specific teams (GS and LAL) with added grouping of their player positions (C and PG).

```{r}
# Find specific stats (PTS, TO) for specific teams (GS, LAL) for specific positions (PG, SG)
nba_team_stats(nba_data, 
               stats_filter = c("PTS","TO"),
               teams_filter = c("GS","LAL"), 
               positions_filter = c("SG","PG"))
```

In the third example, we obtain the descriptive statistics of relevant numeric columns (3PA and FT%) for specific teams (GS) without further grouping of their player positions (NULL input).

```{r}
# Find specific stats (3PA, FT%) for specific teams (GS) without positions_filter inputs
nba_team_stats(nba_data, 
               stats_filter = c("3PA","FT%"), 
               teams_filter = c("GS"))
```

In the fourth example, we obtain the descriptive statistics of relevant numeric columns (3PA and FT%) for all teams (NULL input) with further grouping of specified player positions (PG).

```{r}
# Find specific stats (PTS, TO) for all individual teams (unspecified)
#' # for specific positions (PG, C)
nba_team_stats(nba_data, 
               stats_filter = c("3PA","FT%"),
               positions_filter = c("PG"))
```
